---
title: "Meteorus Sequence Filtering"
author: "NR Graham"
output: html_document
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(readr)
library(tidyverse)
library(seqinr)
library(ggplot2)
```
## Read in the data
```{r readdata, message=FALSE, warning=FALSE}
blast<-read_csv("meteorus_blast.csv")
taxid<-read_csv("meteorus_taxids.csv")
fasta<-read.fasta("metsequences.fasta", as.string = TRUE, forceDNAtolower = FALSE, set.attributes = FALSE)

```
## Explore the data
```{r, exploredata, message=FALSE, warning=FALSE}
names(blast)
names(taxid)
unique(taxid$phylum)
unique(taxid$class)
unique(taxid$order)
```
## Remove non-target sequences
```{r, remove, message=FALSE, warning=FALSE}
met<-blast %>%
  left_join(taxid) %>%
  filter(order == "Hymenoptera") %>%
  select(qseqid, genus, species, pident, length, mismatch, gapopen, evalue, staxids) %>%
  separate(qseqid, c("smpl", "zotu"), remove = FALSE) %>%
  separate(qseqid, c(NA, "size"), sep = ";", remove = FALSE)

length(unique(met$smpl))
 
met %>%
  dplyr::count(genus, staxids) 
  
```
## Subset fasta based on sequences with taxonomic matches to target
```{r, subsetfasta, message=FALSE, warning=FALSE}
subsetlist <- met %>%
  )

subsetlist <- met %>%
  filter(genus == "Meteorus") %>%
  filter(mismatch <= 2) %>%
  filter(length >= 285) %>%
  filter(pident > 99) 

subsetlist %>%
  dplyr::count(genus, species, staxids)

#check to see keeping all samples
length(unique(subsetlist$smpl)) 
  
fasta_filterd<-fasta[names(fasta) %in% subsetlist$qseqid]
```
## Write modified sequences to file
```{r, writefasta, message=FALSE, warning=FALSE}
write.fasta(sequences = fasta_filterd, names = names(fasta_filterd), nbchar = 80, open = "w", file.out = "met_filtered.fasta")
```
## The E-value (expectation value) is a corrected bit-score adjusted to the sequence database size. Lower E-values suggest better hits. May be good to check in some projects.
```{r evalue, echo=TRUE, message=FALSE, warning=FALSE}
sort(unique(met$evalue))
sort(unique(subsetlist$evalue))
```
```{r evalue_hist, echo=FALSE, message=FALSE, warning=FALSE}
met %>%
  ggplot(aes(evalue)) + 
  geom_histogram()+
  theme_minimal()+
  ggtitle("Unfiltered Hymenoptera Hits")

subsetlist %>%
  ggplot(aes(evalue)) + 
  geom_histogram() +
  theme_minimal()+
  ggtitle("Filtered Meteorus spp. Hits")

```

## Playing around with the data to find the best subsetting
```{r}
subsetlist %>%
  filter(species == "(Meteorus ictericus)") %>%
  dplyr::count(smpl, species, zotu) 
subsetlist %>%
  filter(species == "(Meteorus laphygmae)")%>%
  dplyr::count(smpl, species, zotu) 
subsetlist %>%
  filter(species == "(Meteorus pulchricornis)") %>%
  dplyr::count(smpl, species, zotu) 
subsetlist %>%
  filter(staxids == 1031123) %>%
  dplyr::count(smpl, species, zotu) 

```

```{r}

subset2 <- met %>%
  filter(species == "(Meteorus ictericus)" | species == "(Meteorus laphygmae)") %>%
  filter(mismatch <= 2) %>%
  filter(length >= 285) %>%
  filter(pident > 99)

subset2 %>%
  dplyr::count(genus, species, staxids)

#check to see keeping all samples
length(unique(subset2$smpl)) 
  
fasta_filtered2<-fasta[names(fasta) %in% subset2$qseqid]
```

## Write modified sequences to file
```{r, writefasta, message=FALSE, warning=FALSE}
write.fasta(sequences = fasta_filtered2, names = names(fasta_filtered2), nbchar = 80, open = "w", file.out = "met_filtered_laphy89.fasta")
```